\Large Generative Dynamics of Supreme Court Citations: \\ Analysis with a New Statistical Model

### 2020-08-30 Tex/PA_rr/main.tex
revised dependence effects

We include dependence terms to account for each of the dynamics that we hypothesize will characterize the Supreme Court Citation Network---transitivity, reciprocity, and popularity. %We expect that the coefficients associated with each of these dependence effects will be positive.

The common dependence effects for these three dynamics are the number of triangles (case $i$ cites cases $j$ and $k$, and case $j$ cites case $k$), the number of mutual dyads (case $i$ cites $j$ and vice versa), the number of out-2-stars (case $i$ cites $j$ and $k$), and the number of in-2-stars (case $i$ is being cited by cases $j$ and $k$) in $C_{\leq t}$. Unfortunately, adding the triangle, out-2-star, or in-2-star statistic causes model degeneracy \citep{Handcock.2003}. Model degeneracy is a common obstacle when fitting ERGMs, and occurs when the model places most of its probability mass on just a few networks, usually the empty and the full network. In this scenario, the simulated networks are too different from the observed network, making the underlying distribution defined by the model extremely unrealistic.

@@

% popularity
Accounting for the popularity effect through the in-2-star statistic is prone to cause model degeneracy. In an attempt to stabilize the model against model degeneracy as well as to still capture the popularity effect, \citet{SnijdersTomA.B..2006} introduces the \textit{alternating-in-k-star} statistic, which was shown to be equivalent to the \textit{geometrically weighted indegree distribution} (gwidegree) statistic introduced by \citet{hunter2006inference}. Define $D_r(C_t, C_{<t})$ as the number of nodes with indegree $r, r \in \{0, \dots , N_t -1 \}$, where $N_t$ is the number of cases in the network at time $t$, then gwidegree represents a network's indegree distribution in a single statistic by geometrically weighting the degree distribution
\begin{equation}
h_{gwidegree}(C_{t}, C_{<t}, \lambda)= \lambda \sum_{r=1}^{N_t-1}\biggl(1-\bigl(1-\frac{1}{\lambda}\bigr)^r\biggr)D_r(C_t, C_{<t}). 
\label{gwidegree}
\end{equation}
The decay parameter $\lambda \in [0, \infty ]$ controls the decline of the weight put on $D_r(C_t, C_{<t})$ as $r$ increases, which means that the higher $\lambda$ the more the statistic responds to changes for cases with a high indegree. We fixed $\lambda=1$. Adding one single edge to the network changes the indegree distribution of the network so that a case with an indegree of $s$ turns into a case with an indegree of $s + 1$. With our notation this means that $D_r$ and $D_{r+1}$ get replaced by
$D_r -1$ and $D_{r+1} +1$, while no changes for $D_\ell,
 \ell \in \{0, \dots , N_t -1 \} \setminus \{r, r-1\} $ are
made. This changes the log odds in the following way
$$\text{logit} P(C_{ij,t}=1 | C_{-ij,t}, C_{<t}) = \exp \bigl(~\theta \cdot( 1- \frac{1}{\lambda})^r \bigr).$$
As stated by \citet{Levy2016}, this means for the interpretation that a negative parameter value indicates the centralization of edges (i.e., popularity) while a positive parameter indicates the dispersion of edges (i.e., new ties going to less popular nodes). Since we expect highly cited cases to be more likely to be cited again, we expect the gwidegree parameter to be negative.

% Transitivity
We include two statistics to account for different types of transitivity.
The first transitivity statistic calculates the number of transitive ties, $i \to j$, where there is a directed path of length two from $i$ to $j$ through at least one case $k$, and $j$ and $k$ were written during a different term than $i$. We will refer to this statistic as the \textit{different term transitivity} statistic. As illustrated in Section 2, we theorize Supreme Court citations to pattern a high degree of transitivity. As a result, we expect positive parameter estimates for the \textit{different term transitivity} statistic.

The second transitivity statistic, intended to capture clustering that includes more than one tie formed in the same term, is the \textit{geometrically weighted edgewise shared partners} (gwesp) statistic introduced by \citet{hunter2006inference}. For directed networks the edgewise shared $r$-partners statistic ($ESP_r$) can take multiple forms, but we integrate the definition where cases $i$ and $j$ that are connected by a citation, also cite $r$ common other cases. We follow the notation of \citet{Butts.2008} and call this ESP specification \textit{outgoing shared partners} (OSP). Similar to the gwidegree-statistic the gwesp-statistic incorporates a network's edgewise shared partners distribution by replacing $D_r$ by $ESP_r$. We fix the decay parameter at $\lambda=0.25$. Triangles that are incident to the outgoing shared partners configuration arise when there is a cite between two cases decided at the same time, and those two cases cite one or more of the same third cases (decided either previously or also during the same term). This would happen if citations were clustering around legal issues that are not completely captured by our same-issue-area effects.  As we do not expect that the issue area variable will account for all of the clustering around legal issues, we expect the gwesp statistic to also carry a positive coefficient value.

% Activity
We include one more structural network effect to adjust for inherent differences across cases in terms of the overall scope of the legal issues addressed in a case.  \citet{fowler2008authority} finds that cases vary in the degree to which they serve as ``hubs''---citing to many other cases. We expect that new cases will be more likely to cite cases that themselves cited many cases, because cases that themselves cite a large number of cases address a broader array of legal issues. We include the outdegree of every node that entered the network prior time $t$ as a receiver effect. We expect the coefficient on this term to be positive. This is not technically a dependence term, as it is formulated as a covariate, but we discuss it in the current section because it is purely a network structure effect.

>>>>

We include dependence terms to account for each of the dynamics that we hypothesize will characterize the Supreme Court Citation Network---transitivity, reciprocity, and popularity.  The common dependence effects for these three dynamics are the number of triangles, the number of mutual dyads, and the number of in-2-stars. A triangle is a configuration in which there are citations connecting all three cases in a triad. From the perspective of a single tie closing a triad, a positive triangle coefficient indicates that a tie is more likely to form between nodes $i$ and $j$ if nodes $i$ and $j$ have citation ties with one or more cases $k$. A mutual dyad is a configuration in which two nodes within a dyad exchange directed ties. From the perspective of a node closing a mutual dyad with a tie, a positive coefficient indicates that case $i$ is more likely to cite case $j$ if case $j$ has cited case $i$ (mutual dyads are only possible within the same term). The in-2-star configuration is one in which two nodes in a triad send ties to the same third node. From the perspective of a case sending a tie to close an in-2-star, a positive coefficient indicates that case $i$ is more likely to send a tie to case $j$ if one or more other cases also cite case $j$. Unfortunately, adding the triangle or in-2-star statistic causes model degeneracy for the Supreme Court Citation Network---a common issue with these terms \citep{Handcock.2003}. Model degeneracy is a common obstacle when fitting ERGMs, and occurs when the model places most of its probability mass on just a few networks, usually the empty and the full network. In this scenario, the simulated networks are too different from the observed network, making the underlying distribution defined by the model extremely unrealistic.

@@

% popularity
Accounting for the popularity effect through the in-2-star statistic is prone to cause model degeneracy. In an attempt to stabilize the model against model degeneracy as well as to still capture the popularity effect, \citet{SnijdersTomA.B..2006} introduces the \textit{alternating-in-k-star} statistic, which was shown to be equivalent to the \textit{geometrically weighted indegree distribution} (gwidegree) statistic introduced by \citet{hunter2006inference}. The indegree distribution is the distribution of the number of ties sent to nodes, across all nodes in the network. Define $D_r(C_t, C_{<t})$ as the number of nodes with indegree $r, r \in \{0, \dots , N_t -1 \}$, where $N_t$ is the number of cases in the network at time $t$, then gwidegree represents a network's indegree distribution in a single statistic by geometrically weighting the degree distribution
\begin{equation}
h_{gwidegree}(C_{t}, C_{<t}, \lambda)= \lambda \sum_{r=1}^{N_t-1}\biggl(1-\bigl(1-\frac{1}{\lambda}\bigr)^r\biggr)D_r(C_t, C_{<t}). 
\label{gwidegree}
\end{equation}
The decay parameter $\lambda \in [0, \infty ]$ controls the decline of the weight put on each node in the network based on their indegree ($D_r(C_t, C_{<t})$) as the indegree value ($r$ ) increases, which means that the higher $\lambda$ the more the statistic weighs cases with a high indegree. We fixed $\lambda=1$. We chose this value for two reason. First, it results in a fairly straightforward interpretation of the coefficient for gwidegree, as Equation \ref{gwidegree} reduces to a function that counts the number of nodes that receive at least one tie. Our second reason for selecting $\lambda=1$ is that it results in an accurate fit in the indegree distribution of the observed network (as demonstrated in the appendix). At $\lambda=1$, the weight attributed to each case is equivalent regardless of the indegree value, as long as the indegree is greater than zero. This means that the statistic does not grow with the addition of high indegree cases to the network. It only grows through the addition of cases with at least one in-citation. With  $\lambda=1$, the coefficient associated with the gwidegree statistic governs the degree to which the edges are sent to a small number of popular nodes, or distributed fairly evenly across the nodes. A positive coefficient encourages edges to be distributed evenly across a large number of nodes---assuring that as many nodes as possible receive at least one tie. A negative gwidegree coefficient discourages network configurations in which many nodes receive ties, placing higher probability instead on configurations in which many of the ties are sent to a relatively small number of popular nodes. As stated by \citet{Levy2016}, this means for the interpretation that a negative parameter value indicates the centralization of edges (i.e., popularity) while a positive parameter indicates the dispersion of edges (i.e., new ties going to less popular nodes).  Since we expect highly cited cases to be more likely to be cited again, we expect the gwidegree parameter to be negative.

% Transitivity
We include two statistics to account for different types of transitivity. The first transitivity statistic calculates the number of transitive ties, $i \to j$, where there is a directed path of length two from $i$ to $j$ through at least one case $k$, and $j$ and $k$ were written during a different term than $i$. We will refer to this statistic as the \textit{different term transitivity} statistic. This statistic captures the central form of transitivity that we hypothesize above---when a justice writes the opinion for case $i$, cites a past opinion ($k$) as precedent, and then cites one or more of the opinions ($j$) that were cited in opinion $k$. The two-path from opinion $i$ to $j$ is created when opinion $i$ cites an opinion ($k$) that cites an earlier opinion ($j$). The closed transitive triangle is created when opinion $i$ cites directly to $j$, one of the precedents used in opinion $k$. The edge $i \to j$ counts as a ``transitive tie'' since it closes at least one transitive triangle. The \textit{different term transitivity}  statistic is the count of the number of transitive ties in the network that connect opinions written in different terms. As illustrated in Section 2, we theorize Supreme Court citations to exhibit a high degree of transitivity. As a result, we expect positive parameter estimates for the \textit{different term transitivity} statistic.

Whereas the first transitivity statistic is focused on modeling triangles that form through the citation of past cases, the second transitivity statistic we include captures clustering that includes ties formed between cases decided in the same term. We both expect and observe a higher level of transitivity among same-term cases since justices have the opportunity to confer about cases and coordinate citations among related opinions. For modeling within-term transitivity, we use the \textit{geometrically weighted edgewise shared partners} (gwesp) statistic introduced by \citet{hunter2006inference}. We use the form of the gwesp statistic that captures configurations in which cases $i$ and $j$ are connected by a citation ($i$ cites $j$, $j$ cites $i$, or both), and also cite $r$ common other cases (i.e., shared partners). \citet{Butts.2008} call this conceptualization of $r$ the number of \textit{outgoing shared partners}. This form of gwesp captures the degree to which two related cases in the same term are likely to be based on similar bodies of precedent. If the gwesp coefficient is positive, we interpret the effect as saying that the more shared partners $r$ to which cases $i$ and $j$ send ties, the greater the likelihood that there is a citation connecting cases $i$ and $j$.  As with the gwdegree term, there is a decay parameter. The gwesp statistic exhibits decreasing marginal returns with respect to the number of shared partners (e.g., the fourth shared partner between $i$ and $j$ does not contribute as much to gwesp as does the first shared partner between $i$ and $j$). The decay parameter governs the degree of marginal return decrease, with a value of 0 indicating that additional shared partners do not contribute anything, and a value of $\infty$ indicating a linear (non-decreasing) function.  We fix the decay parameter at $\lambda=0.25$, as this accurately captures the edgewise shared partner distribution, and avoids degeneracy (which would be a problem at higher values of $\lambda$). In general, researchers can adjust the gwesp parameter down to avoid degeneracy, and increase it to more accurately capture the shared partner distribution in the network. As with cross-term transitivity, we expect the gwesp statistic to also carry a positive coefficient value.

% Activity
We include one more structural network effect to adjust for inherent differences across cases in terms of the overall scope of the legal issues addressed in a case.  \citet{fowler2008authority} finds that cases vary in the degree to which they serve as ``hubs''---citing to many other cases. We expect that new cases will be more likely to cite cases that themselves cited many cases, because cases that themselves cite a large number of cases address a broader array of legal issues. We include the outdegree (i.e., the number of citations sent) of every node that entered the network prior time $t$ as a receiver effect. We expect the coefficient on this term to be positive, as a positive coefficient would mean that cases that send more citations are likely to also receive more citations. This is not technically a dependence term, as it is formulated as a covariate, but we discuss it in the current section because it is purely a network structure effect.

### 2020-08-30 Tex/PA_rrmemo/PA_RnR_memo.tex
revising memo
revised dependence effects

\noindent \textbf{Addressed:}   Both reviewers' comments on the lack of clarity regarding contributions were very helpful. In our revisions we have both strengthened and clearly stated our substantive and methodological contributions. We offer three contributions---advancing our understanding of Supreme Court citations, illustrating theoretical considerations and methodological choices that should be useful to researchers who study other forms of judicial citations, and a new extension of ERGM that makes it applicable to any sort of citation network. We now state these precisely in both the introduction and conclusion of the paper. To strengthen our substantive contributions, we have expanded the discussion of our results, including integrating that discussing with the literature on Supreme Court citations. To strengthen our methodological contributions, we have expanded our discussion of the c-ERGM itself, and developed a publicly-available R package in which c-ERGM is implemented.

@@

>>>>

\noindent \textbf{Addressed:}  The reviewer's comments on the lack of clarity regarding contributions were very helpful. In our revisions we have both strengthened and clearly stated our substantive and methodological contributions. We offer three contributions---advancing our understanding of Supreme Court citations, illustrating theoretical considerations and methodological choices that should be useful to researchers who study other forms of judicial citations, and a new extension of ERGM that makes it applicable to any sort of citation network. We now state these precisely in both the introduction and conclusion of the paper. To strengthen our substantive contributions, we have expanded the discussion of our results, including integrating that discussing with the literature on Supreme Court citations. To strengthen our methodological contributions, we have expanded our discussion of the c-ERGM itself, and developed a publicly-available R package in which c-ERGM is implemented.

@@

\noindent \textbf{Addressed:}  We agree that the original write-up of the dependence effects fell short in terms of both explaining how to interpret the effects, and providing guidance regarding how we made a number of specification choices. For each dependence effect we revised the coverage in three directions. First, we revised and expanded our explanations of both the intuitive and substantive interpretations of the effect. Second, we provided justifications for functional form and hyperparameter specifications (e.g., the decay parameters in gwidegree and gwesp). Third, we removed technical jargon and mathematical expressions that were not necessary for explaining and justifying what we did.
